
# **Reproducible Research: Peer Assessment 1**

## Loading And Preprocessing the data

Show any code that is needed to

  1. **Load the data(i.e. read.csv())**
  
```{r loading }
setwd("E:/coursera")
##Loading all the packages needed in this exploratory analysis
library(ggplot2)
library(knitr)
library(dplyr)

##Load the data
activity<-read.csv("activity.csv")
##Checking the data
head(activity)
```

```{r}
str(activity)
```

 2. **Process/transform the data (if necessary) into a format suitable for analysis.**

```{r}
##Changing Date from class factor to class date
activity$date<-as.Date(activity$date,format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

  1. **Calculate the total number of steps taken in a day.**
  
```{r totalsteps}
##the total number of steps taken per day is stored in the variable called "total_step"
total_step<-aggregate(steps~date,data=activity,sum,na.rm=TRUE)
head(total_step)
```

 2. **If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day**
 
```{r meanplot}

##Using base plot to make histogram 
hist(total_step$steps,breaks = 20,main = "Total Number of Steps Taken Each Day",
     col = "grey", border = "white", xlab = "Step")
```

 3. **Calculate and report the mean and median of the total number of steps taken per day.**
 
```{r mean}
mean(total_step$steps)
```
```{r}
median(total_step$steps)
```

## What is the average daily activity pattern?

 1. **Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).**
 
```{r}
avg_step<-aggregate(steps~interval,data=activity,mean,na.rm=TRUE)
plot(avg_step$interval, avg_step$steps, type = "l", lwd = 2, col = "navy",
     main = "Time Series: Average Number of Steps Taken", axes = TRUE,
     xlab = "5-minute interval", ylab = "Average number of steps")
```

2. **Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

```{r}
##Find row with maximum number of steps
avg_step_max<-which.max(avg_step$steps)
##Finding interval with this max 
avg_step[avg_step_max,]
```
The 835-th 5-minute interval contains the maximum number of steps.

##Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. **Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).**

```{r missing}
sum(is.na(activity))
```
There are 2304 missing values in the dataset.

2. **Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**

Here I have decided to use the mean of 5-minute interval to fill in the values of the missing values.

3. **Create a new dataset that is equal to the original dataset but with the missing data filled in.**

```{r imputing }
##New Dataset with no missing values is called newactivity.
newactivity<-activity
for (i in avg_step$interval) {
    newactivity[newactivity$interval == i & is.na(newactivity$steps), ]$steps <- 
        avg_step$steps[avg_step$interval == i]
}
head(newactivity) # no NAs
```
```{r}
sum(is.na(newactivity)) ##Should be zero
```

4. **Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?**

```{r imputedplot}
total_step_newactivity<-aggregate(steps~date,data=newactivity,sum,na.rm=TRUE)
hist(total_step_newactivity$steps,breaks = 20,main = "Total Number of Steps Taken Each Day (Imputed)",
     col = "grey", border = "white", xlab = "Step")
```
```{r}
mean(total_step_newactivity$steps)
```
```{r}
median(total_step_newactivity$steps)
```
The mean is the same as the mean from the first part of the assignment, but the median is not, although their values are close. Imputing missing data using the average of the 5-minute interval results in more data points equal to the mean and smaller variation of the distribution. Since many data points have the same values as the mean, the median is much likely to be the same as the mean as well.

##Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

 1. **Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.**
 
```{r}
newactivity$day<-weekdays(newactivity$date)
newactivity$week<-""
newactivity[newactivity$day=="Saturday"|newactivity$day=="Sunday",]$week<-"weekend"
newactivity[!(newactivity$day=="Saturday"|newactivity$day=="Sunday"),]$week<-"weekday"

# convert type_of_day from character to factor
newactivity$week<-factor(newactivity$week)
```
2. **Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.**

```{r}
# calculate average steps by interval across all days
avg_step_newactivity<-aggregate(steps~interval+week,data=newactivity,mean)
# creat a plot
qplot(interval,steps,data=avg_step_newactivity,type="l",geom = c("line"),xlab = "5-minute interval", ylab = "Average number of steps",main = "Average Number of Steps Taken (across all weekday days or weekend days)")+ facet_wrap(~week,ncol=1)
```



